What can GitHub and Googlecode Issues tell us about Exceptions in Android and Java?

### 2014-09-02 android-stacks.tex
all over

In the same rate the number of a users increase, also increases 
the number of users affected by application faults and crashes.
If on the one hand such applications extend phones capabilities 
far beyond of the basic calls and textual messages, on the other hand
they have to face threats to application robustness.
Such applications have to cope with an increasing number of abnormal
computation states, arising from:
 noisy user inputs; faults in underlying middleware or hardware; 
faults on third party services and libraries; compatibility issues [ref]; 
connection with devices e.g., as GPS, camera; memory and battery restrictions; 
 programming errors, among others.

@@

  \item  A multitude of programming mistakes were reported as the main causes of stack traces:
 27,71\% of the analyzed exception stack traces reported a java.lang.NullPointerException the root cause;
and  51,96\% of all projects reported at least one exception stack trace on which the NullPoiterException
was the root cause.

@@

Such categorization was based known sources of faults mobile development: concurrency ~\cite{ama2012}
, backward compatibility  ~\cite{McDon13}, security   ~\cite{enck2011study} and resource management (IO, Memory, Batery) ~\cite{Zhang12}.
We added to this set of categories other three categories: one to represent the exceptions caused by programming mistakes;
  another to represent general Java exceptions (e.g., java.lang.Exception) - whose types do not convey any specific cause of failure; and a category to
represent specific  exceptional situations of apps, frameworks and libraries.
The list of categories is illustrated in Table~\ref{tab:categories}.

@@

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
A multitude of programming mistakes were reported as the main causes of stack traces.
51,96\% of all projects reported at least one exception stack trace on which the NullPoiterException
was the root cause.
}
\end{minipage}}

@@

\emph{\textbf{Runtime Exception wrapping Error.}} We could find  64 exception stack traces 
on 46 projects on which an Error was wrapped in a runtime exception.
Manually inspecting each exception stack trace, we could observe that 44% of the wrappings were performed 
by android.os.AsyncTask. This class is responsible for dealing with asynchronous tasks in Android.
The code snippet is illustrated in Figure ~ref{wrappings} (c) shows that any throwable signaled in the context of an asynchronous 
task is converted in to a runtime and re-thrown. Some of the Erros wrapped on instances of runtime exceptions were 
java.lang.OutOfMemoryError (30), java.lang.NoClassDefFoundError (6), 
java.lang.UnsatisfiedLinkError (5), java.lang.StackOverflowErro (4), as illustrated in Table ~\ref{tab:exampeswrap}

@@

.

@@

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
Odd wrappings were detected (e.g., a checked exception wrapping an OutOfMemoryError)
reveal fragilities on Java Exception Model.
The study also detected an undocumented checked exception signaled by
  a native method defined in Android.
}
\end{minipage}}

@@

\subsection{Inversion of Control and Exception Handling}
Reasoning about the design strategy adopted by the ActivityThread,
we can observe that the inversion of control (on which this class and 
most of the Android framework is based) can make it difficult for the application developer to 
associate a specific handler to the exceptions thrown during the execution of the onCreate method (or any other life cycle method).
According to Java handling model, the handler should be added on a method on the invocation chain. 
However,  in the inversion of control development model the developer does not have fully access to 
the invocation chain of a method. Therefore, the developer may be tempted to convert the checked
 exception into a RuntimeException and resignal.

>>>>

In the same rate the number of a users increase, also increases 
the number of users affected by application faults and crashes.
If on the one hand such applications extend phones capabilities 
far beyond of the basic calls and textual messages, on the other hand
they have to face threats to application robustness ~\cite{was2010}.
Such applications have to cope with an increasing number of abnormal
computation states, arising from:
 noisy user inputs; faults in underlying middleware or hardware; 
faults on third party services and libraries; compatibility issues [ref]; 
connection with devices e.g., as GPS, camera; memory and battery restrictions; 
 programming errors, among others.

@@

  \item  A multitude of programming mistakes were reported as the main causes of exception stack traces:
 27,71\% of the analyzed exception stack traces reported a java.lang.NullPointerException the root cause;
and  51,96\% of all projects reported at least one exception stack trace on which the NullPoiterException
was the root cause.

@@

Such categorization was based known sources of faults mobile development: concurrency ~\cite{ama2012}
, backward compatibility  ~\cite{McDon13}, security   ~\cite{enck2011study,was2010} and resource management (IO, Memory, Batery) ~\cite{Zhang12}.
We added to this set of categories other three categories: one to represent the exceptions caused by programming mistakes;
  another to represent general Java exceptions (e.g., java.lang.Exception) - whose types do not convey any specific cause of failure; and a category to
represent specific  exceptional situations of apps, frameworks and libraries.
The list of categories is illustrated in Table~\ref{tab:categories}.

@@

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
A multitude of programming mistakes were reported as the main causes of exception stack traces.
51,96\% of all projects reported at least one exception stack trace on which the NullPoiterException
was the root cause.
}
\end{minipage}}

@@

\emph{\textbf{Runtime Exception wrapping Error.}} We could find  64 exception stack traces 
on 46 projects on which an Error was wrapped in a runtime exception.
Manually inspecting each exception stack trace, we could observe that 44% of the wrappings were performed 
by android.os.AsyncTask. This class is responsible for dealing with asynchronous tasks in Android.
The code snippet is illustrated in Figure ~ref{wrappings} (c) shows that any throwable signaled in the context of an asynchronous 
task is converted in to a runtime and re-thrown. Some of the Erros wrapped on instances of runtime exceptions were 
java.lang.OutOfMemoryError (30), java.lang.NoClassDefFoundError (6), 
java.lang.UnsatisfiedLinkError (5), java.lang.StackOverflowErro (4), as illustrated in Table ~\ref{tab:exampeswrap}.

@@

@@

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
Odd wrappings were detected (e.g., a checked exception wrapping an OutOfMemoryError).
The study also detected an undocumented checked exception signaled by
  a native method defined in Android.
}
\end{minipage}}

@@

\subsection{Inversion of Control and Exception Handling}

footnote{to do: INPORTANTE Falar que os metodos do ciclo de vida da app nao tem interface excepcional.
desta forma ela não define: O que fazer se o metodo der errado? Nao deixa claro.}

Reasoning about the design strategy adopted by the ActivityThread,
we can observe that the inversion of control (on which this class and 
most of the Android framework is based) can make it difficult for the application developer to 
associate a specific handler to the exceptions thrown during the execution of the onCreate method (or any other life cycle method).
According to Java handling model, the handler should be added on a method on the invocation chain. 
However,  in the inversion of control development model the developer does not have fully access to 
the invocation chain of a method. Therefore, the developer may be tempted to convert the checked
 exception into a RuntimeException and resignal.
